---
title: "Actitity Montoring"
author: "Mary Fowler"
date: "October 1, 2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(tidy.opts= list( width.cutoff = 60), tidy=TRUE)
#install.packages("tidyverse")
library(tidyverse)
library(lubridate)

```

##Task 1: Loading and preprocessing the data

```{r}
#Read in data
activity <- read_csv("activity.csv")
```
##Task 2: Analysis of Steps per Day

For this part of the assignment, you can ignore the missing values in the dataset.

Calculate the total, average and median number of steps taken per day



```{r}
Act_per_day <- activity%>%group_by(date)
steps_per_day <-
  summarise(Act_per_day,
            tot=sum(steps,na.rm=TRUE),
            avg=mean(steps,na.rm=TRUE),
            med=median(steps,na.rm=TRUE))
```

Table of Total, Mean and Median Steps per Day

```{r}
print.data.frame(steps_per_day)
```

Histogram of the total number of steps taken each day

```{r}
ggplot(data= steps_per_day) +
  geom_histogram(mapping=aes(x=tot))
```

##Task 3: Analysis of Average Daily Activity Pattern

```{r}
Act_per_TP <- activity%>%group_by(interval)
steps_per_TP <-
  summarise(Act_per_TP,
            avg=mean(steps,na.rm=TRUE))
```

```{r}
ggplot(data= steps_per_TP, aes(x=interval,y=avg) ) +
  geom_line()
```

```{r}
max_steps_in_TP <- max(steps_per_TP$avg)
 filter(steps_per_TP,max_steps_in_TP==steps_per_TP$avg)

```

Table of Mean Steps per Time Period
```{r}
print.data.frame(steps_per_TP)
```
## Task 4: Impute missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)
```{r}
Num_NA <- sum(is.na(activity$steps))
Num_NA
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

The strategy used in the this analysis: replace the missing steps value with the mean for that time period

Create a new dataset that is equal to the original dataset but with the missing data filled in. dataset = act2

```{r}
activity <- mutate(activity,rn=1:nrow(activity))


act2 <- merge.data.frame(activity, steps_per_TP,
                 by.x="interval",
                 by.y="interval"
                 )
act2[is.na(act2$steps),"steps"] <- act2[is.na(act2$steps),"avg"]
act2 <- act2[order(act2$rn),]
```


Make a histogram of the total number of steps taken each day 

```{r}
Act2_per_day <- act2%>%group_by(date)
steps2_per_day <-
  summarise(Act2_per_day,
            tot=sum(steps,na.rm=TRUE),
            avg=mean(steps,na.rm=TRUE),
            med=median(steps,na.rm=TRUE))
```


```{r}
#steps2_per_day
print.data.frame(steps2_per_day)
```

```{r}
ggplot(data= steps2_per_day) +
  geom_histogram(mapping=aes(x=tot))
```


Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment?

```{r}
cbind(steps_per_day$avg,steps2_per_day$avg,steps_per_day$avg-steps2_per_day$avg)
```

```{r}
cbind(steps_per_day$med,steps2_per_day$med,steps_per_day$med-steps2_per_day$med)
```

```{r}
median(steps_per_TP$avg)
```

What is the impact of imputing missing data on the estimates of the total daily number of steps?

There are fewer time periods with a total steps = 0.

##Task 5: Weekday/Weekend Analysis

Form a factor variable inicating whether a given date is a weekday or weekend day.


```{r}
weekend <- (wday(ymd(act2$date))==1|wday(ymd(act2$date))==7)*1
act2<- cbind(act2,weekend)
act2$weekend <- as.factor(act2$weekend)
act2$weekend <- plyr::mapvalues(act2$weekend, 
                  from=c("0","1"), 
                  to = c("Weekday", "Weekend"))
```

Are there differences in activity patterns between weekdays and weekends?

Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.


```{r}
Act2_per_TP <- act2 %>% group_by(interval,weekend)
steps2_per_TP <-
  summarise(Act2_per_TP,
            avg=mean(steps,na.rm=TRUE))
```

```{r}
ggplot(data= steps2_per_TP, aes(x=interval,y=avg) ) +
  geom_line() +  facet_wrap(~weekend) 


```
















